The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-03-04 paper/paper.tex
add references

\section{Introduction}
Weak gravitational lensing (WL) has emerged over the last decade as a powerful
cosmological probe \citep{cfhtlens,des,kids,hsc}.  WL
uses measurements of coherent shear distortion to the observed shapes of galaxies
to track the evolution of large-scale gravitational fields.  It measures the integrated
gravitational potential along lines of sight to source galaxies, and can thence constrain
the laws of gravity, the expansion history of the Universe, and the history and growth
of cosmic structure.

@@

Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
binning galaxies by redshift.
This approach captures almost all the available information in lensing data, since it measures
an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
gains from moving beyond spectroscopic methods.}.  Binning galaxies by approximate redshift lets us 
model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
While fully 3D methods have been proposed and prototyped \citep{kitching,more}, the tomographic 
approach remains the standard within the field.

Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes \citep{euclid,roman}.

In this paper we discuss and evaluate methods for assigning objects to tomographic bins, and 
specifically the impact
of using a limited set of color bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  Tomographic methods have been 
explored before in \citet{jain} and \citet{kitching2019} \textbf{more}.  In this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux/
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 3\%, far larger than
the requirements for the current generation of surveys \citep{des-y1-cat}.

@@

The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.  It used dark
matter simulations from L-GADGET2 \cite{lgadget} and then added galaxies using
the {\sc AddGals} algorithm \citet{addgals}, which matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
using the LSST band passes were provided in the catalog.

@@

\subsubsection{Metric 1: Sigal-to-noise}
Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a Gaussian estimaate of the covariance between them, as described in
\citet{wl-review} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

@@

By using dilated convolutions, where a fixed step is introduced between adjacent filter taps, the
receptive field of the TCN can be increased without increasing the number of convolutional filters.
Furthermore, residual blocks \citep{he} are also used to stabilize deeper and larger networks. These
modifications cause TCNs to have a longer memory than traditional RNNs with the same capacity,  and
also require low memory for training.

@@

Details of the particular neural network are mostly irrelevant, the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function as to update the parameters $\theta$ of the bin assignment function. This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax2018github}.

>>>>

\section{Introduction}
Weak gravitational lensing (WL) has emerged over the last decade as a powerful
cosmological probe \citep{cfhtlens,rcslens,des,kids,hsc}.  WL
uses measurements of coherent shear distortion to the observed shapes of galaxies
to track the evolution of large-scale gravitational fields.  It measures the integrated
gravitational potential along lines of sight to source galaxies, and can thence constrain
the laws of gravity, the expansion history of the Universe, and the history and growth
of cosmic structure.

@@

Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
binning galaxies by redshift.
This approach captures almost all the available information in lensing data, since it measures
an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
gains from moving beyond spectroscopic methods.}.  Binning galaxies by approximate redshift lets us 
model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
While fully 3D methods have been proposed, prototyped, and shown to have significant promise, 
\citep{heavens,kitching}, the tomographic approach remains the standard within the field.

Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
\citep{euclid,roman}. The general question of optimizing tomography was recently discussed in detail
in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges.

In this paper we discuss and evaluate the related challenge
of using a limited set of color bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  
A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

Furthermore, metacal provides a solution to the pernicious and general problem of
\emph{selection biases} in WL.  Measurements of galaxy shear are known to have noise-induced errors
that are
highly covariant with those on other galaxy properties, including size and, most importantly, flux/
magnitude.  It follows that a cut (or re-weighting) on galaxy magnitudes, or on any quantity derived
from  them, will induce a shear bias since galaxies will preferentially fall either side of the cut
depending on their shear.  Within metacal, we can solve this problem by performing the cut on both
sheared variants of the catalog, and measuring the difference between the post-cut shear in the two
variants.  In DES the corrected biases were found up to around 4\%, far larger than
the requirements for the current generation of surveys \citep{des-y1-cat}.

@@

The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.  It used dark
matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
an algorithm that matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
using the LSST band passes were provided in the catalog.

@@

\subsubsection{Metric 1: Sigal-to-noise}
Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a Gaussian estimaate of the covariance between them, as described in
\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

@@

By using dilated convolutions, where a fixed step is introduced between adjacent filter taps, the
receptive field of the TCN can be increased without increasing the number of convolutional filters.
Furthermore, residual blocks \citep{resnet} are also used to stabilize deeper and larger networks. These
modifications cause TCNs to have a longer memory than traditional RNNs with the same capacity,  and
also require low memory for training.

@@

Details of the particular neural network are mostly irrelevant, the main difficulty in this approach is to be able to compute the back-propagation of gradients through the cosmological loss function as to update the parameters $\theta$ of the bin assignment function. This is was made possible for this challenge by the \texttt{jax-cosmo} library \citep{jax-cosmo}, which allows the computation of both metrics using the JAX automatic differentiation framework \citep{jax}.
